---
title: "Blood Chemistry Analysis \n Buzzard, Goshawk and Red Kite"
author: "Tony Rinaud"
date: "03/03/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, message=F}
library(ggplot2)
library(ggsci)
library(EnvStats)
library(plotly)
library(patchwork)
library(RColorBrewer)
library(chron)
library(RVAideMemoire)

# Routine clean-up upon start
rm(list=ls())
```


```{r data, echo=F}
setwd("C://Users/Tony/Documents/Phd Bielefeld/Data manip/Chemistry_blood/")

chmbl=read.csv2("Chemistry_synlab_2020.csv", h=T, dec=".")

```

```{r str, echo=F}
chmbl$Ring.no=factor(chmbl$Ring.no)
chmbl$RNAseq=factor(chmbl$RNAseq)
chmbl$Lprev=factor(chmbl$Lprev)
chmbl$Alkalische.Phosphatase=as.numeric(chmbl$Alkalische.Phosphatase)
chmbl$AST..GOT.=as.numeric(chmbl$AST..GOT.)
chmbl$LDH=as.numeric(chmbl$LDH)
chmbl$CK..NAC.=as.numeric(chmbl$CK..NAC.)
chmbl$Gesamteiweiï=as.numeric(chmbl$Gesamteiweiï)
chmbl$Date=chron(as.character(chmbl$Date),format="d/m/y")
chmbl$PresLeuco=as.factor(chmbl$PresLeuco)

chmbl=chmbl[order(chmbl$Your_ID,chmbl$Date),]
chmbl=chmbl[!duplicated(chmbl$Ring.no),]
```

# Summary

  * Goshawk has higher Creatine Kinase (CK), Aspartate aminotransferase (AST), Alkalische Phosphatase (AP) and Lactate deshydrogenase (LDH) than Buzzard and Red Kite.
  * Buzzards show high AST and LDH and average CK when infected, indicative of liver cell damage.


# Results blood chemistry

## Gallensäuren: 

### Comparison infected / non-infected

**Common Buzzard**

  * Significantly higher in infected nestlings than non-infected 
  
  * But no differences between classes of infection intensity

```{r}

# Comparison infected / non-infected birds

# Two Sample t-test  
#  
#  data:  buzz_Gallensauren by buzz_PresLeuco  
#  t = -2.5693, df = 65, p-value = 0.0125  
#  alternative hypothesis: true difference in means is not equal to 0  
#  95 percent confidence interval:  
#   -14.018445  -1.756376  
#  sample estimates:  
#  mean in group 0 mean in group 1   
#         24.81571        32.70312 


# Comparison of classes of infection intensity

# Permutation Analysis of Variance Table
# 
# Response: buzz$Gallensauren
# 999 permutations
#             Sum Sq Df Mean Sq F value Pr(>F)
# buzz$Lprev  1118.9  4  279.73  1.7068  0.165
# Residuals  10161.2 62  163.89 
```

```{r compmean gall buzz, echo=F}

buzz=chmbl[chmbl$Species %in% "buzz",]

# qqPlot(buzz$Gallensauren[buzz$Lprev %in% c(1,2,3,4)])
# qqPlot(buzz$Gallensauren[buzz$Lprev %in% c(0)])
# hist(buzz$Gallensauren[buzz$Lprev %in% c(1,2,3,4)])
# hist(buzz$Gallensauren[buzz$Lprev %in% c(0)])

# Homogeneity of variances

# var.test(buzz$Gallensauren~buzz$PresLeuco)  # homogeneous variances 

# Mean comparison

# t.test(buzz$Gallensauren~buzz$PresLeuco, var.equal=T) # Difference between Infected non infected

```

**Goshawk**

  Can't compare between non infected/infected Goshawk, two low sample sizes.

```{r compmean gall gos, echo=F}

gos=chmbl[chmbl$Species %in% "gos",]

meansgall=aggregate(Gallensauren~Species, chmbl, mean)

ggplot(chmbl)+
  aes(Species, Gallensauren)+
  geom_boxplot(aes(fill=Species))+
  stat_n_text(text.box = 1)+
  theme_minimal()+
  stat_summary(fun.y=mean, colour="darkred", geom="point", 
               shape=18, size=3,show.legend = FALSE)+
  geom_label(data = meansgall, aes(label = round(Gallensauren,digits = 0), x=c(1.1,2.1,3.1), y = Gallensauren), size=2.9)

```

**Red Kite**

  * Higher but non significant difference of Gallensauren in infected nestlings vs non-infected
  * No difference between classes of infection intensity
  
```{r}

# Comparison of infected and non-infected birds

# Permutation Two Sample t-test  
#  
#  data:  rk_Gallensauren by rk_PresLeuco  
#  999 permutations  
#  t = -1.4334, p-value = 0.136  
#  alternative hypothesis: true difference in means is not equal to 0   
#  sample estimates:  
#  **mean in group 0 mean in group 1**  
#         **36.59412**        **46.79583** 

# Comparison among classes of infection intensity 

# Permutation Analysis of Variance Table
# 
# Response: rk$Gallensauren
# 999 permutations
#           Sum Sq Df Mean Sq F value Pr(>F)
# rk$Lprev   957.6  3  319.18  0.8493  0.492
# Residuals 9395.5 25  375.82
```
        
```{r compmean gall rk, echo=F, message=F, warning=F}

rk=chmbl[chmbl$Species %in% "rk",]

# qqPlot(rk$Gallensauren[rk$Lprev %in% c(1,2,3,4)])
# qqPlot(rk$Gallensauren[rk$Lprev %in% c(0)])
# hist(rk$Gallensauren[rk$Lprev %in% c(1,2,3,4)])
# hist(rk$Gallensauren[rk$Lprev %in% c(0)])

# Homogeneity of variances

# var.test(rk$Gallensauren~rk$PresLeuco)  # homogeneous variances 

# Mean comparison

# perm.t.test(rk$Gallensauren~rk$PresLeuco) # No difference between Infected non infected in Red Kite

```

## GLDH: 

### Comparison infected / non-infected

* Most of the samples are in the class <1.0: not possible to compare between groups

## Creatine Kinase (CK) 

**Common Buzzard**

* Mean is way over 400 (significant, mean = 2282), the reference value.

* No difference between infected / non infected

```{r}

# Comparison mean samples and theorical value (400) 

# One Sample t-test
# 
# data:  buzz$CK..NAC.
# t = 19.062, df = 66, p-value < 2.2e-16
# alternative hypothesis: true mean is not equal to 400
# 95 percent confidence interval:
#  2085.001 2479.267
# sample estimates:
# mean of x 
#  2282.134 

# Comparison infected - non infected

# Two Sample t-test  
#    
#  data:  buzz_CK..NAC. by buzz_PresLeuco  
#  t = -0.50878, df = 65, p-value = 0.6126  
#  alternative hypothesis: true difference in means is not equal to 0  
#  95 percent confidence interval:  
#   -498.1510  295.8706  
#  sample estimates:  
#  **mean in group 0 mean in group 1**  
#         **2233.829**  **2334.969** 
```

```{r compmean ck buzz, echo=F}

# Comparaison avec valeur théorique = 400 

# t.test(buzz$CK..NAC.,mu=400)

# qqPlot(buzz$CK..NAC.[buzz$Lprev %in% c(1,2,3,4)])
# qqPlot(buzz$CK..NAC.[buzz$Lprev %in% c(0)])
# hist(buzz$CK..NAC.[buzz$Lprev %in% c(1,2,3,4)])
# hist(buzz$CK..NAC.[buzz$Lprev %in% c(0)])

# Homogeneity of variances

# var.test(buzz$CK..NAC.~buzz$PresLeuco)  # homogeneous variances 

# Mean comparison

# t.test(buzz$CK..NAC.~buzz$PresLeuco, var.equal=T) # No difference between Infected non infected

```

**Goshawk**

* Same, but the 4 points are higher than Buzzard's and Red Kite's 

* Ck higher in Goshawk than Buzzard (non-sign.) and Red Kite(significant, p=0.003). No differences between Buzzard and Goshawk.

```{r compmean ck gos, echo=F}

meanCK=aggregate(CK..NAC.~Species,chmbl,mean)

ggplot(chmbl)+
  aes(Species,CK..NAC.)+
  geom_boxplot(aes(fill=Species))+
  stat_n_text(text.box = 1)+
  stat_summary(fun.y=mean, colour="darkred", geom="point", 
               shape=18, size=3,show.legend = FALSE)+
  geom_label(data = meanCK, aes(label = round(CK..NAC.,digits = 0), x=c(1.2,2.2,3.2), y = CK..NAC.), size=2.9)

```

```{r}
# bartlett.test(chmbl$CK..NAC.~chmbl$Species)
# 
# perm.anova(chmbl$CK..NAC.~chmbl$Species)
# 
# pairwise.perm.t.test(chmbl$CK..NAC.,chmbl$Species)
```

**Red Kite**

  * Over the reference, 400 (sign., mean rk around 1462)
  
  * No difference between infected and non infected nestlings

```{r}

# Comparaison avec valeur théorique = 400 

# One Sample t-test
# 
# data:  rk$CK..NAC.
# t = 9.7061, df = 28, p-value = 1.852e-10
# alternative hypothesis: true mean is not equal to 400
# 95 percent confidence interval:
#  1238.172 1686.587
# sample estimates:
# mean of x 
#  1462.379 

# Comparison infected / non-infected birds

# Two Sample t-test  
#  
#  data:  rk_CK..NAC. by rk_PresLeuco  
#  t = 0.13984, df = 27, p-value = 0.8898  
#  alternative hypothesis: true difference in means is not equal to 0  
#  95 percent confidence interval:  
#   -432.5551  495.8296  
#  sample estimates:  
#  **mean in group 0 mean in group 1**   
#         **1475.471**        **1443.833**
```
        
```{r compmean ck rk, echo=F}

# t.test(rk$CK..NAC.,mu=400)

# qqPlot(rk$CK..NAC.[rk$Lprev %in% c(1,2,3,4)])
# qqPlot(rk$CK..NAC.[rk$Lprev %in% c(0)])
# hist(rk$CK..NAC.[rk$Lprev %in% c(1,2,3,4)])
# hist(rk$CK..NAC.[rk$Lprev %in% c(0)])

# Homogeneity of variances

# var.test(rk$CK..NAC.~rk$PresLeuco)  # homogeneous variances 

# Mean comparison

# t.test(rk$CK..NAC.~rk$PresLeuco, var.equal=T) # No difference between Infected non infected
```

## Aspartate aminotransferase (AST)

**Common Buzzard**

* Mean is within the range of reference (100-350) i.e 293.

* Difference between infected / non infected AST

```{r}

# Comparison mean versus reference value (225) 

# One Sample t-test
# 
# data:  buzz$AST..GOT.
# t = 3.6706, df = 66, p-value = 0.0004858
# alternative hypothesis: true mean is not equal to 225
# 95 percent confidence interval:
#  256.2095 330.6562
# sample estimates:
# mean of x 
#  293.4328

# Comparison infect / non-infected birds

# Permutation Two Sample t-test
# 
# data:  buzz_AST..GOT. by buzz_PresLeuco  
# 999 permutations  
# t = -3.1274, p-value = 0.004  
# alternative hypothesis: true difference in means is not equal to 0  
# sample estimates:  
# **mean in group 0 mean in group 1**   
#        **241.0571**        **350.7188** 
```

```{r compmean AST buzz, echo=F}

## Comparaison avec valeur théorique = 400

# t.test(buzz$AST..GOT.,mu=mean(c(100,350)))
# 
# qqPlot(buzz$AST..GOT.[buzz$Lprev %in% c(1,2,3,4)])
# qqPlot(buzz$AST..GOT.[buzz$Lprev %in% c(0)])
# hist(buzz$AST..GOT.[buzz$Lprev %in% c(1,2,3,4)])
# hist(buzz$AST..GOT.[buzz$Lprev %in% c(0)])
# 
# ## Homogeneity of variances
# 
# var.test(buzz$AST..GOT.~buzz$PresLeuco)  # homogeneous variances
# 
# ## Mean comparison
# 
# perm.t.test(buzz$AST..GOT.~buzz$PresLeuco) # No difference between Infected non infected

```

**Goshawk**

* Same, but those 4 points are higher than Buzzard's and Red Kite's

* Difference between Goshawk and Buzzard; Red kite and Buzzard. But not between Red kite and Goshawk.

```{r}
# Pairwise comparisons using permutation t tests 
# 
# data:  chmbl$AST..GOT. and chmbl$Species
# 999 permutations 
# 
#     buzz  gos  
# gos 0.087 -             # p-value
# rk  0.087 0.168         # p-values
# 
# P value adjustment method: fdr 
```

```{r compmean AST gos, echo=F}

meansAST <- aggregate(AST..GOT. ~  Species, chmbl, mean)

ggplot(chmbl)+
  aes(Species,AST..GOT.)+
  geom_boxplot(aes(fill=Species))+
  stat_n_text(text.box = 1)+
  stat_summary(fun.y=mean, colour="darkred", geom="point", 
               shape=18, size=3,show.legend = FALSE)+
  geom_label(data = meansAST, aes(label = round(AST..GOT.,digits = 0), x=c(1.1,2.1,3.1), y = AST..GOT.), size=2.9)

```

```{r, echo=F}
# bartlett.test(chmbl$AST..GOT.~chmbl$Species)
# 
# perm.anova(chmbl$AST..GOT.~chmbl$Species)
# 
# pairwise.perm.t.test(chmbl$AST..GOT.,chmbl$Species)
```

**Red Kite**

  * Over 400 (mean around 1450)
  
  * No difference between infected and non infected nestlings

```{r}
#  Welch Two Sample t-test  
# 
# data:  rk_AST..GOT. by rk_PresLeuco  
# t = -0.69244, df = 15.128, p-value = 0.4992  
# alternative hypothesis: true difference in means is not equal to 0  
# 95 percent confidence interval:  
#  -156.02373   79.46491  
# sample estimates:  
# **mean in group 0 mean in group 1**   
#        **350.4706**        **388.7500** 
```
  
  
```{r compmean AST rk, echo=F}
# qqPlot(log(rk$AST..GOT.[rk$Lprev %in% c(1,2,3,4)]))
# qqPlot(log(rk$AST..GOT.[rk$Lprev %in% c(0)]))
# hist(log(rk$AST..GOT.[rk$Lprev %in% c(1,2,3,4)]))
# hist(log(rk$AST..GOT.[rk$Lprev %in% c(0)]))
# 
# # Homogeneity of variances
# 
# var.test(log(rk$AST..GOT.)~rk$PresLeuco)  # homogeneous variances
# 
# # Mean comparison
# 
# t.test(rk$AST..GOT.~rk$PresLeuco, var.equal=F) # No difference between Infected non infected
```

## Lactate Deshydrogenase (LDH)

**Common Buzzard**

* Mean above the range of reference (150-450) i.e 1003.418.

* Difference between infected / non infected

```{r}

# Comparison rk LDH mean and reference value = mean(log(value)) to correct for normality and heterogeneity of variances 

# One Sample t-test
# 
# data:  log(buzz$LDH)
# t = 17.546, df = 66, p-value < 2.2e-16
# alternative hypothesis: true mean is not equal to 5.703782
# 95 percent confidence interval:
#  6.650974 6.894209
# sample estimates:
# mean of x 
#  6.772592 
 
# Comparison Infected / non-infected
 
# Welch Two Sample t-test
# 
# data:  log(buzz_LDH) by buzz_PresLeuco  
# t = -3.6083, df = 61.112, p-value = 0.0006222   
# alternative hypothesis: true difference in means is not equal to 0  
# 95 percent confidence interval:  
#  -0.6328719 -0.1815562  
# sample estimates:  
# **mean in group 0 mean in group 1**   
#        **6.578102**        **6.985316**
```
       
       ==> AST and LDH higher and CK same as non-infected in infected: indicative of liver cell damage (cf In Practice March 2015 Vol 37)


```{r compmean LDH buzz, echo=F}

## Comparaison avec valeur théorique = 400
# 
# t.test(log(buzz$LDH),mu=log(mean(c(150,450))))
# 
# qqPlot(log(buzz$LDH[buzz$Lprev %in% c(1,2,3,4)]))
# qqPlot(log(buzz$LDH[buzz$Lprev %in% c(0)]))
# hist(log(buzz$LDH[buzz$Lprev %in% c(1,2,3,4)]))
# hist(log(buzz$LDH[buzz$Lprev %in% c(0)]))
# 
# ## Homogeneity of variances
# 
# var.test(log(buzz$LDH)~buzz$PresLeuco)  # homogeneous variances
# 
# ## Mean comparison
# 
# t.test(log(buzz$LDH)~buzz$PresLeuco) # Difference between Infected non infected

```

**Goshawk**

* Same, but those 4 points are higher than Buzzard's and Red Kite's 

* All different between each species with Goshawk being higher than two others

```{r}
# Pairwise comparisons using permutation t tests 
# 
# data:  chmbl$LDH and chmbl$Species
# 999 permutations 
# 
#     buzz  gos  
# gos 0.027 -    
# rk  0.042 0.006
# 
# P value adjustment method: fdr 
```


```{r compmean LDH gos, echo=F}

meansLDH <- aggregate(LDH ~  Species, chmbl, mean)

ggplot(chmbl)+
  aes(Species,LDH)+
  geom_boxplot(aes(fill=Species))+
  stat_n_text(text.box = 1)+
  stat_summary(fun.y=mean, colour="darkred", geom="point", 
               shape=18, size=3,show.legend = FALSE)+
  geom_label(data = meansLDH, aes(label = round(LDH,digits = 0), x=c(1.20,2.20,3.20), y = LDH + 50))

```

```{r, echo=F}

# bartlett.test(chmbl$LDH~chmbl$Species)
# 
# t.test(c(buzz$LDH,rk$LDH)~c(buzz$Species,rk$Species))
# 
# pairwise.perm.t.test(chmbl$LDH,chmbl$Species)

```

**Red Kite**

  * Over 400 (mean around 1450)
  
  * No difference between infected and non infected nestlings
 
```{r}
#  Welch Two Sample t-test  
# 
# data:  rk_LDH by rk_PresLeuco  
# t = 1.1278, df = 25.35, p-value = 0.27  
# alternative hypothesis: true difference in means is not equal to 0  
# 95 percent confidence interval:  
#  -110.3592  377.9278  
# sample estimates:  
# **mean in group 0 mean in group 1**   
#        **804.1176**        **670.3333**
```
 
  
```{r compmean LDH rk, echo=F}
# qqPlot(rk$LDH[rk$Lprev %in% c(1,2,3,4)])
# qqPlot(rk$LDH[rk$Lprev %in% c(0)])
# hist(rk$LDH[rk$Lprev %in% c(1,2,3,4)])
# hist(rk$LDH[rk$Lprev %in% c(0)])
# 
# # Homogeneity of variances
# 
# var.test(rk$LDH~rk$PresLeuco)  # homogeneous variances
# 
# # Mean comparison
# 
# t.test(rk$LDH~rk$PresLeuco, var.equal=F) # No difference between Infected non infected
```

## Alkalische Phosphatase (AP)

**Common Buzzard**

* Mean above the range of reference (10-250) i.e 413.

* No difference between infected / non infected

```{r}

# Comparison mean versus reference value (130)

# One Sample t-test
# 
# data:  buzz$Alkalische.Phosphatase
# t = 14.67, df = 66, p-value < 2.2e-16
# alternative hypothesis: true mean is not equal to 130
# 95 percent confidence interval:
#  374.1090 451.0253
# sample estimates:
# mean of x 
#  412.5672 

# Comparison infected / non-infected nestlings

# Welch Two Sample t-test
# 
# data:  buzz$Alkalische.Phosphatase by buzz$PresLeuco
# t = 1.3397, df = 61.44, p-value = 0.1853
# alternative hypothesis: true difference in means is not equal to 0
# 95 percent confidence interval:
#  -24.92057 126.15629
# sample estimates:
# mean in group 0 mean in group 1 
#        436.7429        386.1250
```


```{r compmean Alkalische.Phosphatase buzz, echo=F}

## Comparaison avec valeur théorique = 130

# t.test(buzz$Alkalische.Phosphatase,mu=mean(c(10,250)))
# 
# qqPlot(buzz$Alkalische.Phosphatase[buzz$Lprev %in% c(1,2,3,4)])
# qqPlot(buzz$Alkalische.Phosphatase[buzz$Lprev %in% c(0)])
# hist(buzz$Alkalische.Phosphatase[buzz$Lprev %in% c(1,2,3,4)])
# hist(buzz$Alkalische.Phosphatase[buzz$Lprev %in% c(0)])
# 
# ## Homogeneity of variances
# 
# var.test(buzz$Alkalische.Phosphatase~buzz$PresLeuco)  # homogeneous variances
# 
# ## Mean comparison
# 
# t.test(buzz$Alkalische.Phosphatase~buzz$PresLeuco) # Difference between Infected non infected

```

**Goshawk**

Same, but the 4 points are higher than Buzzard's and Red Kite's 

```{r compmean AP gos, echo=F}

meanAP=aggregate(Alkalische.Phosphatase~Species, chmbl, mean)

ggplot(chmbl)+
  aes(Species,Alkalische.Phosphatase)+
  geom_boxplot(aes(fill=Species))+
  stat_n_text(text.box = 1)+
  stat_summary(fun.y=mean, colour="darkred", geom="point", 
               shape=18, size=3,show.legend = FALSE)+
  geom_label(data = meanAP, aes(label = round(Alkalische.Phosphatase,digits = 0), x=c(1.20,2.20,3.20), y = Alkalische.Phosphatase + 50))

```

**Red Kite**

  * Over 250 (mean around 356)
  
  * No difference between infected and non infected nestlings

```{r}
# Comparison mean rk versus reference value (130)

# One Sample t-test
# 
# data:  rk$Alkalische.Phosphatase
# t = 15.8, df = 28, p-value = 1.769e-15
# alternative hypothesis: true mean is not equal to 130
# 95 percent confidence interval:
#  327.4496 386.2745
# sample estimates:
# mean of x 
#  356.8621 

# Comparison infected / non-infected nestlings
 
#   Welch Two Sample t-test  
# 
# data:  rk_Alkalische.Phosphatase by rk_PresLeuco   
# t = 1.6996, df = 26.044, p-value = 0.1011  
# alternative hypothesis: true difference in means is not equal to 0  
# 95 percent confidence interval:  
#  -9.233957 97.469251  
# sample estimates:  
# **mean in group 0 mean in group 1**   
#        **375.1176**        **331.0000**
```


```{r compmean AP rk, echo=F}

# t.test(rk$Alkalische.Phosphatase,mu=mean(c(10,250)))

# qqPlot(rk$Alkalische.Phosphatase[rk$Lprev %in% c(1,2,3,4)])
# qqPlot(rk$Alkalische.Phosphatase[rk$Lprev %in% c(0)])
# hist(rk$Alkalische.Phosphatase[rk$Lprev %in% c(1,2,3,4)])
# hist(rk$Alkalische.Phosphatase[rk$Lprev %in% c(0)])
# 
# # Homogeneity of variances
# 
# var.test(rk$Alkalische.Phosphatase~rk$PresLeuco)  # homogeneous variances
# 
# # Mean comparison
# 
# t.test(rk$Alkalische.Phosphatase~rk$PresLeuco, var.equal=F) # No difference between Infected non infected
```

## Gamma glutamyl transferase (g.GT)

**Common Buzzard**

* No difference between infected / non infected or between classes of infection

```{r}
# Comparison classes of infection intensity

# Permutation Analysis of Variance Table
# 
# Response: 1/log(buzz$g.GT)
# 999 permutations
#             Sum Sq Df   Mean Sq F value Pr(>F)
# buzz$Lprev 0.03531  4 0.0088282  0.6089  0.672
# Residuals  0.89888 62 0.0144980

# Comparison infected / non-infected

# Permutation Two Sample t-test  
# 
# data:  1/log(buzz_g.GT) by buzz_PresLeuco  
# 999 permutations  
# t = 0.2415, p-value = 0.828  
# alternative hypothesis: true difference in means is not equal to 0  
# sample estimates:  
# **mean in group 0 mean in group 1**   
#       **0.5586918**      **0.5516137** 
```


```{r compmean g.GT buzz, echo=F}

## Comparaison avec valeur théorique = 400

# qqPlot(1/buzz$g.GT[buzz$Lprev %in% c(1,2,3,4)])
# qqPlot(1/buzz$g.GT[buzz$Lprev %in% c(0)])
# hist(1/log(buzz$g.GT[buzz$Lprev %in% c(1,2,3,4)]))
# hist(1/log(buzz$g.GT[buzz$Lprev %in% 0]))
# 
# ## Homogeneity of variances
# 
# var.test(1/log(buzz$g.GT)~buzz$PresLeuco)  # homogeneous variances
# 
# ## Mean comparison
# 
# perm.t.test(1/log(buzz$g.GT)~buzz$PresLeuco) # No difference between Infected non infected

```

**Goshawk**

Same, but the 4 points are lower than Buzzard's and Red Kite's 

```{r compmean g.GT gos, echo=F}

meansgGT=aggregate(g.GT~Species, chmbl, mean)

ggplot(chmbl)+
  aes(Species,g.GT)+
  geom_boxplot(aes(fill=Species))+
  stat_n_text(text.box = 1)+
  stat_summary(fun.y=mean, colour="darkred", geom="point", 
               shape=18, size=3,show.legend = FALSE)+
  geom_label(data = meansgGT, aes(label = round(g.GT,digits = 1), x=c(1.20,2.20,3.20), y = g.GT))

```

**Red Kite**
  
  * No difference between infected and non infected nestlings and classes of infection
  
```{r}

# Comparison of classes of infection

# Permutation Analysis of Variance Table
# 
# Response: 1/log(rk$g.GT)
# 999 permutations
#            Sum Sq Df  Mean Sq F value Pr(>F)
# rk$Lprev  0.10837  3 0.036124  1.9579  0.103
# Residuals 0.46126 25 0.018450

# Comparison infected / non-infected nestlings

# Permutation Two Sample t-test  
# 
# data:  1/rk_g.GT by rk_PresLeuco  
# 999 permutations  
# t = 1.7234, p-value = 0.106  
# alternative hypothesis: true difference in means is not equal to 0  
# sample estimates:  
# **mean in group 0 mean in group 1**   
#       **0.1928359**       **0.1461826**
```


```{r compmean g.GT rk, echo=F}
# qqPlot(1/rk$g.GT[rk$Lprev %in% c(1,2,3,4)])
# qqPlot(1/rk$g.GT[rk$Lprev %in% c(0)])
# hist(1/rk$g.GT[rk$Lprev %in% c(1,2,3,4)])
# hist(1/rk$g.GT[rk$Lprev %in% c(0)])
# 
# # Homogeneity of variances
# 
# var.test(1/rk$g.GT~rk$PresLeuco)  # homogeneous variances
# 
# # Mean comparison
# 
# perm.t.test(1/rk$g.GT~rk$PresLeuco) # No difference between Infected non infected
```

## Albumin

* Could biased by heparin tubes...

**Common Buzzard**

* No difference between infected / non infected

```{r}
# Two Sample t-test  
# 
# data:  buzz_Albumin by buzz_PresLeuco  
# t = -0.47737, df = 65, p-value = 0.6347  
# alternative hypothesis: true difference in means is not equal to 0  
# 95 percent confidence interval:  
#  -1.3523823  0.8305966  
# sample estimates:  
# **mean in group 0 mean in group 1**   
#        **9.582857**        **9.843750** 
```

  

```{r compmean Albumin buzz, echo=F}

## Comparaison avec valeur théorique = 400

# qqPlot(buzz$Albumin[buzz$Lprev %in% c(1,2,3,4)])
# qqPlot(buzz$Albumin[buzz$Lprev %in% c(0)])
# hist(buzz$Albumin[buzz$Lprev %in% c(1,2,3,4)])
# hist(buzz$Albumin[buzz$Lprev %in% 0])
# 
# ## Homogeneity of variances
# 
# var.test(buzz$Albumin~buzz$PresLeuco)  # homogeneous variances
# 
# ## Mean comparison
# 
# t.test(buzz$Albumin~buzz$PresLeuco, var.equal=T) # Difference between Infected non infected

```

**Goshawk**

Same, but the 4 points are higher (but spread) than Buzzard's and Red Kite's 

```{r compmean Albumin gos, echo=F}

meansalb=aggregate(Albumin~Species,chmbl,mean)

ggplot(chmbl)+
  aes(Species,Albumin)+
  geom_boxplot(aes(fill=Species))+
  stat_n_text(text.box = 1)+
  stat_summary(fun.y=mean, colour="darkred", geom="point", 
               shape=18, size=3,show.legend = FALSE)+
  geom_label(data = meansalb, aes(label = round(Albumin,digits = 1), x=c(1.20,2.20,3.20), y = Albumin))

```

**Red Kite**
  
  * No difference between infected and non infected nestlings

```{r}
# Two Sample t-test
# data:  rk$Albumin by rk$PresLeuco  
# t = 1.28, df = 27, p-value = 0.2114  
# alternative hypothesis: true difference in means is not equal to 0  
# 95 percent confidence interval:  
#  -0.8252831  3.5625380  
# sample estimates:  
# **mean in group 0 mean in group 1**   
#        **12.03529**        **10.66667**
```

   

```{r compmean Albumin rk, echo=F}

# qqPlot(rk$Albumin[rk$Lprev %in% c(1,2,3,4)])
# qqPlot(rk$Albumin[rk$Lprev %in% c(0)])
# hist(rk$Albumin[rk$Lprev %in% c(1,2,3,4)])
# hist(rk$Albumin[rk$Lprev %in% c(0)])
# 
# # Homogeneity of variances
# 
# var.test(rk$Albumin~rk$PresLeuco)  # homogeneous variances
# 
# # Mean comparison
# 
# t.test(rk$Albumin~rk$PresLeuco, var.equal=T) # No difference between Infected non infected
```

# PCA

```{r}

library(ade4)
library(ggcorrplot)
library(vegan)
require(FactoMineR)
require(factoextra)
require(ggplot2)
require(tidyr)
require(dplyr)
require(MASS)
require(reshape2)
require(cowplot)
library(patchwork)
library(EnvStats)
library(ggrepel)

setwd("C://Users/Tony/Documents/Phd Bielefeld/Data manip/Chemistry_blood/")

bloody=read.csv2('raw_07.07.20.csv', h=T, dec=',')

# Infection classes No / Low / High Infection

bloody$infectionclass=NA

for (i in 1:length(bloody$Infection_intensity)){
  if(bloody$Infection_intensity[i]==0){bloody$infectionclass[i]="No infection"}
  if(bloody$Infection_intensity[i] %in% c(1,2)){bloody$infectionclass[i]="Low infection"}
  if(bloody$Infection_intensity[i] %in% c(3,4)){bloody$infectionclass[i]="High infection"}
} 

bloodypca=bloody[,c(10:16,19:28)]

bloodypca=scale(bloodypca)

bloodypca=na.omit(bloodypca)

# Correlation plot between explanatory variables

ggcorrplot(cor(bloodypca), p.mat = cor_pmat(bloodypca), hc.order=TRUE, type='lower')

# Function circular plot

circleFun <- function(center = c(0,0),diameter = 1, npoints = 100){
  r = diameter / 2
  tt <- seq(0,2*pi,length.out = npoints)
  xx <- center[1] + r * cos(tt)
  yy <- center[2] + r * sin(tt)
  return(data.frame(x = xx, y = yy))
}

circ <- circleFun(c(0,0),2,npoints = 500)

# PCA  (cf https://tem11010.github.io/Plotting-PCAs/)

PCAblood=PCA(bloodypca)

plot.PCA(PCAblood)

pc1=unlist(PCAblood$ind$coord[, 1], recursive = T,use.names = T)
pc2=unlist(PCAblood$ind$coord[, 2], recursive = T,use.names = T)

coords=data.frame(pc1,pc2)

pca.vars <- PCAblood$var$coord %>% data.frame

pca.vars$vars <- rownames(pca.vars)

pca.vars.m <- melt(pca.vars, id.vars = "vars")

# Plot PCA variables correlation

vars.p <-  ggplot() +
               geom_path(data = circ,aes(x,y), lty = 2, color = "grey", alpha = 0.7)+
               geom_hline(yintercept = 0, lty = 2, color = "grey", alpha = 0.9)+
               geom_vline(xintercept = 0, lty = 2, color = "grey", alpha = 0.9)+
               geom_segment(data = pca.vars, aes(x = 0, xend = Dim.1, y = 0, yend = Dim.2),
                            arrow = arrow(length = unit(0.025, "npc"), type = "open"), 
                            lwd = 1)+ 
              geom_text_repel(data = pca.vars, 
                        aes(x = Dim.1*1.15, y =  Dim.2*1.15, 
                            label = c("Leukocyte_count","Lymphocytes_percent","Monocytes_percent","Heterophiles_percent","H_L_ratio","Basophiles_percent","Eosinophiles_percent","AP","AST","gGT","CHE","GLDH","BA","LDH","CK","ALB","TP")), 
                            check_overlap = F, size = 5)+
              xlab("PC 1")+ 
              ylab("PC2")+
              coord_equal()+
              theme_minimal()+
              theme(panel.grid = element_blank(), 
                    panel.border = element_rect(fill= "transparent"))

vars.p

# Plot PCA individual with species labels

indplot=ggplot(data=coords, aes(x = pc1, y = pc2, color=bloody$Species))+
  geom_point(color='black',size=4.5)+
  geom_point(size=4)+
  labs(x="PC1 (16.89%)", y="PC2 (14.98%)",color = "Species")+
  geom_hline(yintercept = 0, lty = 2)+
  geom_vline(xintercept = 0, lty = 2)+
  scale_color_manual(values=c("#427EF6", "#7C3918", "#FF9100"))+
  theme_classic()+
  theme(legend.position = "bottom", legend.text = element_text(size=12),axis.title=element_text(size=14), axis.text=element_text(size=12), plot.title = element_text(hjust = 0.5, face="bold"), panel.border = element_rect(fill= "transparent"))+
  stat_ellipse(geom="polygon", aes(fill = bloody$Species),alpha = 0.2, show.legend = FALSE, level = 0.95)

########################
# Plot PCA individuals with infection intensity labels 

indinfplot=ggplot(data=coords, aes(x = pc1, y = pc2, color=as.factor(bloody$infectionclass))) +
  geom_point(color='black',size=4.5)+
  geom_point(size=4)+
  labs(x="PC1 (16.89%)", y="PC2 (14.98%)",color = "Infection intensity")+
  geom_hline(yintercept = 0, lty = 2)+
  geom_vline(xintercept = 0, lty = 2)+
  scale_color_manual(values=c("#A50F15","#FB6A4A","#FCAE91"))+
  theme_classic()+
  theme(legend.position = "bottom", legend.text = element_text(size=12),axis.title=element_text(size=14), axis.text=element_text(size=12), plot.title = element_text(hjust = 0.5, face="bold"), panel.border = element_rect(fill= "transparent"))+
  stat_ellipse(geom="polygon", aes(color = as.factor(bloody$infectionclass)), alpha = 0.05, show.legend = FALSE, level = 0.95, size=0.7)

indplot/indinfplot|vars.p + plot_layout(ncol=1)

```


# LDA

```{r}

#LDA constrained by Species

qqPlot(bloodypca[,1])

anova(betadisper(dist(bloodypca),bloody$Species))

ldaSpecies=lda(bloodypca, bloody$Species)   #  CV=T ajout script Olivier

#tableclass=base::table(catptPGR, ldaPGR$class)

MVA.cv(bloodypca, bloody$Species,model="LDA")

```

```{r}
# Group means:
#                      Leukocyte_count Lymphocytes_percent Monocytes_percent
# Y.Milvus milvus           -0.1013353           0.1087629        -0.9984180
# Y.Accipiter gentilis      -0.2259058           0.1487041        -0.4635512
# Y.Buteo buteo              0.4489071          -0.3060109         1.1187630
#                      Heterophiles_percent  H_L_ratio Basophiles_percent
# Y.Milvus milvus                -0.4310962 -0.2182354         0.25040583
# Y.Accipiter gentilis           -0.2565877 -0.2208657        -0.09105667
# Y.Buteo buteo                   0.5888760  0.4686820         0.10812979
#                      Eosinophiles_percent          AP        AST        gGT        CHE
# Y.Milvus milvus                 0.4502106  1.18532884  1.4004485  0.1677907  0.7358960
# Y.Accipiter gentilis            0.1157404 -0.03055676 -0.3762194 -0.1045012  0.3663715
# Y.Buteo buteo                  -0.3295659 -0.23903828  0.3552992  0.1539921 -0.8709206
#                             GLDH         BA         LDH         CK        ALB
# Y.Milvus milvus       0.61803236 -0.2333693  1.88885581  1.2406125  0.9079120
# Y.Accipiter gentilis -0.02621955 -0.2496934 -0.07435103  0.1732512 -0.2695516
# Y.Buteo buteo        -0.10534642  0.5265174 -0.33280578 -0.6349992  0.2784312
#                              TP
# Y.Milvus milvus       0.8917281
# Y.Accipiter gentilis -0.2928062
# Y.Buteo buteo         0.3260797

``` 
```{r}
par(mfrow=c(1,2))
MVA.plot(ldaSpecies,fac=bloody$Species,xlab="Axis 1", ylab="Axis 2", col=c("#427EF6", "#7C3918", "#FF9100"),legend=F, legend.cex=0.5, legend.pos="topleft", legend.lab=c("Accipiter gentilis", "Buteo buteo","Milvus milvus"))
MVA.plot(ldaSpecies,"corr", xlab="Axis 1", ylab="Axis 2")

#labels=c("Response\nrate","Latency", "Number\n songs","Score"))

# mod=lm(as.matrix(bvPGR)~catptPGR)
# 
# plotresid(mod)
# 
# Manova(mod, test.statistic="Wilks")
# 
# summary(mod)

```

```{r}

#LDA constrained by Infection class

qqPlot(bloodypca[,1])

anova(betadisper(dist(bloodypca),bloody$infectionclass))

ldaInfclass=lda(bloodypca, bloody$infectionclass)   #  CV=T ajout script Olivier

#tableclass=base::table(catptPGR, ldaPGR$class)

# MVA.cv(bloodypca, bloody$infectionclass,model="LDA")

par(mfrow=c(1,2))
MVA.plot(ldaInfclass,fac=bloody$infectionclass,xlab="Axis 1", ylab="Axis 2", col=c("#427EF6", "#7C3918", "#FF9100"),legend=F, legend.cex=0.9, legend.pos="topleft", legend.lab=c("High Infection", "Low Infection","No Infection"))
MVA.plot(ldaInfclass,"corr", xlab="Axis 1", ylab="Axis 2")

# mod=lm(as.matrix(bvPGR)~catptPGR)
# 
# plotresid(mod)
# 
# Manova(mod, test.statistic="Wilks")
# 
# summary(mod)
```

